Addition and removal energies of circular quantum dots

### 2017-06-13 Manuscript/paper.tex
Reflow paragraphs and clean up whitespace

Even though the model system contains five  parameters $(N, m, \omega,
e, \epsilon)$, many of them are redundant.  With an appropriate choice
of units, the number of parameters in the system can be reduced to
just two.  For rest of this paper, we shall use atomic
units
%\footnote{Here, ``atomic unit'' is used in an \emph{abstract}
%sense: we do not assume (nor care) whether $m$, $e$, $\epsilon$
%correspond to true values in vacuum, or some effective value in
%medium.} 
where $\hbar$, $m$, $e$ and $4 \pi \epsilon$ are all set to
one.  The unit Hartree $E_{\mathrm{h}}$ becomes the unit of energy and the Bohr
radius $a$ becomes the unit of length:
\begin{align*}
  E_{\mathrm{h}} &\equiv m \left(\frac{e^2}{4 \pi \epsilon \hbar}\right)^2 &
  a &\equiv \frac{4 \pi \epsilon \hbar^2}{m e^2}.
\end{align*}
This leaves us with $(N, \omega)$ as the only parameters necessary to
specify the quantum dot system, with $\omega$ in units of
$E_{\mathrm{h}} / \hbar$.

@@

In second quantization, the one- and two-body Hamiltonian operators $H^{[1]}$ and $\hat{H}^{[2]}$ from Eq.~\eqref{eq:onetwobodyhamiltonian} can be rewritten as
\begin{align} \label{eq:second_quantized_hamiltonian}
  \hat H^{[1]} &= \sum_{pq} \langle p | \hat{H}^{[1]} | q \rangle \hat a_p^\dagger \hat a_q^{}, &
  \hat{H}^{[2]} &= \frac{1}{4} \sum_{pqrs} \langle pq | \hat{H}^{[2]} | rs \rangle \hat a_p^\dagger \hat a_q^\dagger \hat a_s^{} \hat a_r^{}.
\end{align}
where the quantities $\langle p | \hat{H}^{[1]} | q \rangle$
and $\langle pq | \hat{H}^{[2]} | rs \rangle$ are known
as \textit{matrix elements} and fully characterize the two operators
in the basis.  In this paper, we use the letters $p, q, r, s,
t, \ldots$ to label arbitrary single-particle \textit{states} in the
basis.  For the harmonic oscillator basis, labels such as $p$ can be
considered abbreviations for triplets of the form $(n, m_\ell,
m_{\mathrm{s}})$.  A summation such as $\sum_p$ is then understood as
a summation over every single-particle basis state $p$.  This is not
to be confused with summations such as $\sum_j$
from \eqref{eq:onetwobodyhamiltonian}, which contains a label $j$
denoting an arbitrary \textit{particle}.  We thus see another benefit
with second quantization: the Hamiltonian representation makes no
attempt to distinguish particles at all, as expected for particles
with quantum statistics.

The complete Hamiltonian is given by the sum
\begin{align}
  \hat{H} = \sum_{p,q} \langle p | \hat{H}^{[1]} | q \rangle \hat{a}_p^\dagger \hat{a}_q + \frac{1}{4} \sum_{pqrs} \langle pq | \hat{H}^{[2]} | rs \rangle \hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r . \label{eq:vacuumhamiltonian}
\end{align}
Due to the antisymmetry of Slater determinants, we have the property
that exchanging either $p$ with $q$ or $r$ with $s$ reverses the sign
of $\langle pq | \hat{H}^{[2]} | rs \rangle$, leading to a 4-fold
symmetry that must be canceled out by the $1/4$ factor in front of the
summation in \eqref{eq:vacuumhamiltonian}.  For this reason, the
two-body matrix element $\langle pq | \hat{H}^{[2]} | rs \rangle$
is commonly referred to as the \emph{antisymmetrized} matrix element.
This is not to be confused with the (non-antisymmetric) interaction
integral,
\begin{align} \label{eq:interactionintegral}
  v_{p_1p_2p_3p_4} \equiv \int \langle p_1 | \bm{r} \rangle \langle p_2 | \bm{r}' \rangle v(\bm r, \bm r') \langle \bm{r} | p_3 \rangle \langle \bm{r}' | p_4 \rangle \D \bm{r} \D \bm{r}'.
\end{align}
This object only has a two-fold symmetry: $v_{p_1p_2p_3p_4} = v_{p_2p_1p_4p_3}$.  The antisymmetrized matrix element are related to the interaction integral via
\begin{align} \label{eq:antisymmetricmatrixelement}
  \langle pq | \hat{H}^{[2]} | rs \rangle = v_{pqrs} - v_{pqsr}
\end{align}

@@

To perform numerical calculations, we further assume that each unknown
state $|q'\rangle$ is built from a linear combination of known states
$|p\rangle$, with an unknown matrix of coefficients $\bm C$ defined
via the transformation equation
\begin{align*}
  |q'\rangle \equiv \sum_p |p\rangle C_{pq'}.
\end{align*}
This allows the problem to be reduced from an abstract minimization
problem \eqref{eq:hfexpectation} to a concrete numerical problem.  The
caveat is that the set of known functions be large enough to capture
the relevant behavior of the system.

@@

The goal then is to find the coefficients $\bm C$ that minimize the energy $E_{\Phi}$, now given by
\begin{align}
  E_{\Phi} &= \sum_i n_i H^{[1] \prime}_{ii} + \sum_{ij} n_i n_j H^{[2] \prime}_{ijij} \label{eq:hfenergy},
\end{align}
where the transformed operators are defined as
\begin{align}
  H^{[1] \prime}_{pq} &\equiv \sum_{rs} C_{rp}^* \langle r | \hat{H}^{[1]} | s \rangle C_{sq}^{} \label{eq:hftransform1} \\
  H^{[2] \prime}_{pqrs} &\equiv \sum_{tuvw} C_{tp}^* C_{uq}^* \langle tu | \hat{H}^{[2]} | vw \rangle C_{vr}^{} C_{ws}^{}, \label{eq:hftransform2}
\end{align}
and $n_p$ is the \textit{occupancy}, defined as
\begin{align}
  n_p \equiv \begin{cases}
    1 & \text{if $p$ is an occupied state in $\Phi$}, \\
    0 & \text{if $p$ is an unoccupied state in $\Phi$}.
  \end{cases}
\end{align}
We will also use $\bar n_p \equiv 1 - n_p$ as a shorthand for the complement.

With the method of Lagrange multipliers, the minimization problem can
be reduced to the solving of a nonlinear equation --
the \textit{Hartree--Fock equation}:
\begin{align} \label{eq:hartreefock}
  \bm{F} \bm{C} = \bm{C} \bm{\varepsilon},
\end{align}
where the \textit{Fock matrix} $\bm F$ is defined as
\begin{align} \label{eq:fock}
  F_{pq} \equiv \langle p | \hat{H}^{[1]} | q \rangle + \sum_{rsi} n_i C_{ri}^* \langle pr | \hat{H}^{[2]} | qs \rangle C_{si}^{},
\end{align}
and $\bm{\varepsilon}$ is a vector of Lagrange multipliers: each element $\varepsilon_{q'}$ is associated with a specific single-particle state $|q'\rangle$.

@@

The second-quantized operators in
Eq.~\eqref{eq:second_quantized_hamiltonian} are constructed with
respect to the vacuum state in the sense that their operator strings
$\hat{a}_p^\dagger \hat{a}_q$ and
$\hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r$ both have
expectation values of zero in vacuum:
\begin{align*}
&\bra{} \hat{a}_p^\dagger \hat{a}_q \ket{} = 0, &
&\bra{} \hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r \ket{} = 0.
\end{align*}
They are said to be \emph{normal-ordered} with respect to the vacuum state $\ket{}$.  It is possible to be normal-ordered with respect to a different state, however.  Suppose we have a Slater determinant $\Phi$ with $N$ occupied states:
\begin{align*}
  \ket{\Phi} \equiv \ket{i_1, i_2, \ldots, i_N}
\end{align*}
where $\{i_1, i_2, \ldots, i_N\}$ is a set of $N$ indices that label
the occupied states.  One can define a set of \textit{quasiparticle
field operators} $\hat b$ and $\hat b^\dagger$ that obey the same
anticommutation relations as the $\hat a$ operators.  The normal
ordering of the quasiparticle field operators is straightforward to
compute.  What is more interesting is the normal ordering of the
physical field operators $\hat a$ and $\hat a^\dagger$ with respect to
$|\Phi\rangle$:
\begin{align*}
  \normord{\hat{a}_p^\dagger \hat{a}_q^{}} &=
  \hat{a}_p^\dagger \hat{a}_q^{} - n_p^{} \delta_{p, q}^{}.
\end{align*}
The normal ordering notation thus provides a concise way of expressing the right-hand side.  In this new notation, we can rewrite the Hamiltonian $\hat H = \hat{H}^{[1]} + \hat{H}^{[2]}$ in its \emph{normal-ordered representation}:
\begin{align}
  \hat{H} = H^{[0]} + \sum_{pq} H^{[1]}_{pq} \normord{\hat{a}_p^\dagger \hat{a}_q^{}} + \frac{1}{4} \sum_{pqrs} H^{[2]}_{pqrs} \normord{\hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s^{} \hat{a}_r^{}},
  \label{eq:normordhamiltonian}
\end{align}
where
\begin{align*}
  H^{[0]} &\equiv \sum_{i} n_i \langle i | \hat{H}^{[1]} | i \rangle + \frac{1}{2} \sum_{ij} n_i n_j \langle ij | \hat{H}^{[2]} | ij \rangle, \\
  H^{[1]}_{pq} &\equiv \langle p | \hat{H}^{[1]} | q \rangle + \sum_{i} n_i \langle pi | \hat{H}^{[2]} | qi \rangle, \\
  H^{[2]}_{pqrs} &\equiv \langle pq | \hat{H}^{[2]} | rs \rangle. \\
\end{align*}
If $\Phi$ is a state optimized by the HF method, then $H^{[0]}$ is
simply the HF energy $E_\Phi$ and $H^{[1]}_{pq}$ is the Fock matrix
$F_{pq}$.  The operator $\hat H$ in
Eq.~\eqref{eq:normordhamiltonian} is completely equivalent to $\hat H$
in Eq.~\eqref{eq:vacuumhamiltonian}.  The only difference is that the
meaning of a ``$k$-body operator'' has been redefined with respect to
the reference state $\Phi$ rather than vacuum state $\ket{}$, causing
matrix elements to be reshuffled among the $k$-body operators.  Where
this makes a critical difference is when the operator expressions
are \emph{truncated}, i.e.\ some $k$-body operators may be discarded
from the computation for efficiency reasons.  Normal-ordering can
drastically alter the importance of the $k$-body operators with
respect to each other.

@@

With the operator truncation, the generator $\hat{\eta}$ can be written as a generic 2-body operator:
\begin{align*}
\hat{\eta} = \sum_{pq} \eta_{pq}^{[1]} \normord{\hat a_p^\dagger \hat a_q} +
\frac{1}{4} \sum_{pqrs}\eta_{pqrs}^{[2]} \normord{\hat a_p^\dagger \hat a_q^\dagger \hat a_s \hat a_r},
\end{align*}
where $\eta_{pq}^{[1]}$ and $ \eta_{pqrs}^{[2]}$ respectively are its
one- and two-body matrix elements normal ordered with respect to
$\Phi$, subject to the antihermittivity constraint.

By expanding the commutator in Eq.~\eqref{eq:imsrgode} and discarding
the three-body term, we obtain the matrix-element form of the
IM-SRG(2) flow equation:
\begin{align}
    \frac{\D H^{[k]}}{\D s} = C^{[k]}(\eta, H) \label{eq:flowmxe}
\end{align}
where
\begin{align}
  C^{[k]}(A, B) &\equiv 2 \mathcal{A}_{A, B} D^{[k]}(A, B) \text{ for any $k$, $A$, $B$}, \label{eq:commutmxe} \\
  \mathcal{A}_{x, y} f(x, y) &\equiv \frac{1}{2} \bigl(f(x, y) - f(y, x)\bigr) \text{ for any $x$, $y$, $f$},
\end{align}
\begin{align}
  D^{[0]}(A, B)
  &\equiv
    \sum_{ia} n_i \bar{n}_a A^{[1]}_{ia} B^{[1]}_{ai}
    + \frac{1}{4} \sum_{ijab} n_i n_j \bar{n}_a \bar{n}_b A^{[2]}_{ijab} B^{[2]}_{abij},
    \label{eq:flow0} \\
  D^{[1]}_{p, q}(A, B)
  &\equiv
    \sum_{r} A^{[1]}_{pr} B^{[1]}_{rq}
    - \frac{1}{2} \sum_{ija} n_i n_j \bar{n}_a A^{[2]}_{ijaq} B^{[2]}_{apij}
    + \frac{1}{2} \sum_{iab} n_i \bar{n}_a \bar{n}_b A^{[2]}_{ipab} B^{[2]}_{abiq}
    \notag \\
  &\hphantom{=}
    + \sum_{ia} n_i \bar{n}_a \left(
    A^{[1]}_{ia} B^{[2]}_{apiq}
    + A^{[2]}_{ipaq} B^{[1]}_{ai}
    \right)
    \label{eq:flow1} \\
  D^{[2]}_{pqrs}(A, B)
  &\equiv
    \tilde{\mathcal{A}}_{p, q} \tilde{\mathcal{A}}_{rs}
    \sum_{ia} n_i \bar{n}_a A^{[2]}_{ipar} B^{[2]}_{aqis}
    \notag \\
  &\hphantom{=}
    + \frac{1}{2} \sum_{ij} n_i n_j A^{[2]}_{ijrs} B^{[2]}_{pqij}
    + \frac{1}{2} \sum_{ab} \bar{n}_a \bar{n}_b A^{[2]}_{pqab} B^{[2]}_{abrs}
    \notag \\
  &\hphantom{=}
    + \sum_{t} \left(
    \tilde{\mathcal{A}}_{pq} A^{[1]}_{qt} B^{[2]}_{ptrs}
    + \tilde{\mathcal{A}}_{rs} A^{[2]}_{pqrt} B^{[1]}_{ts}.
    \right)
    \label{eq:flow2}
\end{align}

@@

The White generator is explicitly constructed the following way \cite{PhysRevLett.106.222502,White:cond-mat0201346}.  Let
\begin{align*}
\hat{\eta}^{\text{Wh}} &\equiv \hat{\eta}' - \hat{\eta}'{}^\dagger,
\end{align*}
where
\begin{align*}
\eta^{\prime [1]}_{ai} &\equiv \frac{\bar{n}_a n_i H^{[1]}_{ai}}{\tilde{\Delta}^{[1]}_{ai}} &
\eta^{\prime [2]}_{abij} &\equiv \frac{\bar{n}_a \bar{n}_b n_i n_j H^{[2]}_{abij}}{\tilde{\Delta}^{[2]}_{abij}}
\end{align*}
and the Epstein--Nesbet denominators $\tilde{\Delta}$ are defined as {\bf ref to epstein-nesbet deno, insert later}
\begin{align}
\tilde{\Delta}^{[1]}_{ai} &\equiv \Delta_{\{a\}, \{i\}} - H^{[2]}_{aiai}, \notag \\
\tilde{\Delta}^{[2]}_{abij} &\equiv \Delta_{\{a, b\}, \{i, j\}} + w_{abij}, \notag \\
\Delta^{[2]}_{P_1P_2} &\equiv \sum_{p_1 \in P_1} H^{[1]}_{p_1p_1} - \sum_{p_2 \in P_2} H^{[1]}_{p_2p_2}, \label{eq:moellerplessetdenominator} \\
w_{abij}
  &\equiv H^{[2]}_{abab} - H^{[2]}_{aiai} - H^{[2]}_{bibi} \notag \\
  &\qquad + H^{[2]}_{ijij} - H^{[2]}_{ajaj} - H^{[2]}_{bjbj}. \notag
\end{align}
Some authors {\bf ref}  use the M\o ller--Plesset denominators $\Delta$
instead, leading to a slightly different variant of the White
generator.
{\bf define Moller plesset denos and give reference}.

@@

We can then express the perturbation expansion in terms of summations
over matrix elements as we did for the IM-SRG flow
equation, again with the aid of Wick's theorem or diagrammatic
techniques.  This leads to the following expression for the
second-order correction:
\begin{align*}
  \varepsilon_p^{(2)}
  &=
    \sum_{iab} \frac{n_i \bar{n}_a \bar{n}_b |H^{[2]}_{piab}|^2}{2 \Delta_{\{pi\}, \{ab\}}}
    - \sum_{ija} \frac{n_i n_j \bar{n}_a |H^{[2]}_{ijpa}|^2}{2 \Delta_{\{ij\}, \{pa\}}},
\end{align*}
where the M\o ller--Plesset denominators $\Delta$ are defined in
Eq.~\eqref{eq:moellerplessetdenominator}.  The above expression is
depicted in Figure \ref{fig:diagrams-sfe}.  Since there are numerous
terms in the third-order correction $\varepsilon_p^{(3)}$, they are
listed in diagrammatic form in Figure \ref{fig:diagrams-sfe}.  We will
refer to QDPT to third order as ``QDPT3''.

>>>>

Even though the model system contains five parameters $(N, m, \omega, e, \epsilon)$, many of them are redundant.  With an appropriate choice of units, the number of parameters in the system can be reduced to just two.  For rest of this paper, we shall use atomic units where $\hbar$, $m$, $e$ and $4 \pi \epsilon$ are all set to one.  The unit Hartree $E_{\mathrm{h}}$ becomes the unit of energy and the Bohr radius $a$ becomes the unit of length:
\begin{align*}
  E_{\mathrm{h}} &\equiv m \left(\frac{e^2}{4 \pi \epsilon \hbar}\right)^2 &
  a &\equiv \frac{4 \pi \epsilon \hbar^2}{m e^2}.
\end{align*}
This leaves us with $(N, \omega)$ as the only parameters necessary to specify the quantum dot system, with $\omega$ in units of $E_{\mathrm{h}} / \hbar$.

@@

In second quantization, the one- and two-body Hamiltonian operators $H^{[1]}$ and $\hat{H}^{[2]}$ from Eq.~\eqref{eq:onetwobodyhamiltonian} can be rewritten as
\begin{align} \label{eq:second_quantized_hamiltonian}
  \hat H^{[1]} &= \sum_{p q} \langle p | \hat{H}^{[1]} | q \rangle \hat a_p^\dagger \hat a_q^{}, &
  \hat{H}^{[2]} &= \frac{1}{4} \sum_{p q r s} \langle p q | \hat{H}^{[2]} | r s \rangle \hat a_p^\dagger \hat a_q^\dagger \hat a_s^{} \hat a_r^{}.
\end{align}
where the quantities $\langle p | \hat{H}^{[1]} | q \rangle$ and $\langle p q | \hat{H}^{[2]} | r s \rangle$ are known as \textit{matrix elements} and fully characterize the two operators in the basis.  In this paper, we use the letters $p, q, r, s, t, \ldots$ to label arbitrary single-particle \textit{states} in the basis.  For the harmonic oscillator basis, labels such as $p$ can be considered abbreviations for triplets of the form $(n, m_\ell, m_{\mathrm{s}})$.  A summation such as $\sum_p$ is then understood as a summation over every single-particle basis state $p$.  This is not to be confused with summations such as $\sum_j$ from \eqref{eq:onetwobodyhamiltonian}, which contains a label $j$ denoting an arbitrary \textit{particle}.  We thus see another benefit with second quantization: the Hamiltonian representation makes no attempt to distinguish particles at all, as expected for particles with quantum statistics.

The complete Hamiltonian is given by the sum
\begin{align}
  \hat{H} = \sum_{p, q} \langle p | \hat{H}^{[1]} | q \rangle \hat{a}_p^\dagger \hat{a}_q + \frac{1}{4} \sum_{p q r s} \langle p q | \hat{H}^{[2]} | r s \rangle \hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r. \label{eq:vacuumhamiltonian}
\end{align}
Due to the antisymmetry of Slater determinants, we have the property that exchanging either $p$ with $q$ or $r$ with $s$ reverses the sign of $\langle p q | \hat{H}^{[2]} | r s \rangle$, leading to a 4-fold symmetry that must be canceled out by the $1/4$ factor in front of the summation in \eqref{eq:vacuumhamiltonian}.  For this reason, the two-body matrix element $\langle p q | \hat{H}^{[2]} | r s \rangle$ is commonly referred to as the \emph{antisymmetrized} matrix element.  This is not to be confused with the (non-antisymmetric) interaction integral,
\begin{align} \label{eq:interactionintegral}
  v_{p_1 p_2 p_3 p_4} \equiv \int \langle p_1 | \bm{r} \rangle \langle p_2 | \bm{r}' \rangle v(\bm r, \bm r') \langle \bm{r} | p_3 \rangle \langle \bm{r}' | p_4 \rangle \D \bm{r} \D \bm{r}'.
\end{align}
This object only has a two-fold symmetry: $v_{p_1 p_2 p_3 p_4} = v_{p_2 p_1 p_4 p_3}$.  The antisymmetrized matrix element are related to the interaction integral via
\begin{align} \label{eq:antisymmetricmatrixelement}
  \langle p q | \hat{H}^{[2]} | r s \rangle = v_{p q r s} - v_{p q s r}
\end{align}

@@

To perform numerical calculations, we further assume that each unknown state $|q'\rangle$ is built from a linear combination of known states $|p\rangle$, with an unknown matrix of coefficients $\bm C$ defined via the transformation equation
\begin{align*}
  |q'\rangle \equiv \sum_p |p\rangle C_{p q'}.
\end{align*}
This allows the problem to be reduced from an abstract minimization problem \eqref{eq:hfexpectation} to a concrete numerical problem.  The caveat is that the set of known functions be large enough to capture the relevant behavior of the system.

@@

The goal then is to find the coefficients $\bm C$ that minimize the energy $E_{\Phi}$, now given by
\begin{align}
  E_{\Phi} &= \sum_i n_i H^{[1] \prime}_{i i} + \sum_{i j} n_i n_j H^{[2] \prime}_{i j i j} \label{eq:hfenergy},
\end{align}
where the transformed operators are defined as
\begin{align}
  H^{[1] \prime}_{p q} &\equiv \sum_{r s} C_{r p}^* \langle r | \hat{H}^{[1]} | s \rangle C_{s q}^{} \label{eq:hftransform1} \\
  H^{[2] \prime}_{p q r s} &\equiv \sum_{t u v w} C_{t p}^* C_{u q}^* \langle t u | \hat{H}^{[2]} | v w \rangle C_{v r}^{} C_{w s}^{}, \label{eq:hftransform2}
\end{align}
and $n_p$ is the \textit{occupancy}, defined as
\begin{align}
  n_p \equiv \begin{cases}
    1 & \text{if $p$ is an occupied state in $\Phi$}, \\
    0 & \text{if $p$ is an unoccupied state in $\Phi$}.
  \end{cases}
\end{align}
We will also use $\bar n_p \equiv 1 - n_p$ as a shorthand for the complement.

With the method of Lagrange multipliers, the minimization problem can be reduced to the solving of a nonlinear equation -- the \textit{Hartree--Fock equation}:
\begin{align} \label{eq:hartreefock}
  \bm{F} \bm{C} = \bm{C} \bm{\varepsilon},
\end{align}
where the \textit{Fock matrix} $\bm F$ is defined as
\begin{align} \label{eq:fock}
  F_{p q} \equiv \langle p | \hat{H}^{[1]} | q \rangle + \sum_{r s i} n_i C_{r i}^* \langle p r | \hat{H}^{[2]} | q s \rangle C_{s i}^{},
\end{align}
and $\bm{\varepsilon}$ is a vector of Lagrange multipliers: each element $\varepsilon_{q'}$ is associated with a specific single-particle state $|q'\rangle$.

@@

The second-quantized operators in Eq.~\eqref{eq:second_quantized_hamiltonian} are constructed with respect to the vacuum state in the sense that their operator strings $\hat{a}_p^\dagger \hat{a}_q$ and $\hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r$ both have expectation values of zero in vacuum:
\begin{align*}
&\bra{} \hat{a}_p^\dagger \hat{a}_q \ket{} = 0, &
&\bra{} \hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r \ket{} = 0.
\end{align*}
They are said to be \emph{normal-ordered} with respect to the vacuum state $\ket{}$.  It is possible to be normal-ordered with respect to a different state, however.  Suppose we have a Slater determinant $\Phi$ with $N$ occupied states:
\begin{align*}
  \ket{\Phi} \equiv \ket{i_1, i_2, \ldots, i_N}
\end{align*}
where $\{i_1, i_2, \ldots, i_N\}$ is a set of $N$ indices that label the occupied states.  One can define a set of \textit{quasiparticle field operators} $\hat b$ and $\hat b^\dagger$ that obey the same anticommutation relations as the $\hat a$ operators.  The normal ordering of the quasiparticle field operators is straightforward to compute.  What is more interesting is the normal ordering of the physical field operators $\hat a$ and $\hat a^\dagger$ with respect to $|\Phi\rangle$:
\begin{align*}
  \normord{\hat{a}_p^\dagger \hat{a}_q^{}} &=
  \hat{a}_p^\dagger \hat{a}_q^{} - n_p^{} \delta_{p, q}^{}.
\end{align*}
The normal ordering notation thus provides a concise way of expressing the right-hand side.  In this new notation, we can rewrite the Hamiltonian $\hat H = \hat{H}^{[1]} + \hat{H}^{[2]}$ in its \emph{normal-ordered representation}:
\begin{align}
  \hat{H} = H^{[0]} + \sum_{p q} H^{[1]}_{p q} \normord{\hat{a}_p^\dagger \hat{a}_q^{}} + \frac{1}{4} \sum_{p q r s} H^{[2]}_{p q r s} \normord{\hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s^{} \hat{a}_r^{}},
  \label{eq:normordhamiltonian}
\end{align}
where
\begin{align*}
  H^{[0]} &\equiv \sum_{i} n_i \langle i | \hat{H}^{[1]} | i \rangle + \frac{1}{2} \sum_{i j} n_i n_j \langle i j | \hat{H}^{[2]} | i j \rangle, \\
  H^{[1]}_{p q} &\equiv \langle p | \hat{H}^{[1]} | q \rangle + \sum_{i} n_i \langle p i | \hat{H}^{[2]} | q i \rangle, \\
  H^{[2]}_{p q r s} &\equiv \langle p q | \hat{H}^{[2]} | r s \rangle. \\
\end{align*}
If $\Phi$ is a state optimized by the HF method, then $H^{[0]}$ is simply the HF energy $E_\Phi$ and $H^{[1]}_{p q}$ is the Fock matrix $F_{p q}$.  The operator $\hat H$ in Eq.~\eqref{eq:normordhamiltonian} is completely equivalent to $\hat H$ in Eq.~\eqref{eq:vacuumhamiltonian}.  The only difference is that the meaning of a ``$k$-body operator'' has been redefined with respect to the reference state $\Phi$ rather than vacuum state $\ket{}$, causing matrix elements to be reshuffled among the $k$-body operators.  Where this makes a critical difference is when the operator expressions are \emph{truncated}, i.e.\ some $k$-body operators may be discarded from the computation for efficiency reasons.  Normal-ordering can drastically alter the importance of the $k$-body operators with respect to each other.

@@

With the operator truncation, the generator $\hat{\eta}$ can be written as a generic 2-body operator:
\begin{align*}
\hat{\eta} = \sum_{p q} \eta_{p q}^{[1]} \normord{\hat a_p^\dagger \hat a_q} +
\frac{1}{4} \sum_{p q r s}\eta_{p q r s}^{[2]} \normord{\hat a_p^\dagger \hat a_q^\dagger \hat a_s \hat a_r},
\end{align*}
where $\eta_{p q}^{[1]}$ and $ \eta_{p q r s}^{[2]}$ respectively are its one- and two-body matrix elements normal ordered with respect to $\Phi$, subject to the antihermittivity constraint.

By expanding the commutator in Eq.~\eqref{eq:imsrgode} and discarding the three-body term, we obtain the matrix-element form of the IM-SRG(2) flow equation:
\begin{align}
    \frac{\D H^{[k]}}{\D s} = C^{[k]}(\eta, H) \label{eq:flowmxe}
\end{align}
where
\begin{align}
  C^{[k]}(A, B) &\equiv 2 \mathcal{A}_{A, B} D^{[k]}(A, B) \text{ for any $k$, $A$, $B$}, \label{eq:commutmxe} \\
  \mathcal{A}_{x, y} f(x, y) &\equiv \frac{1}{2} \bigl(f(x, y) - f(y, x)\bigr) \text{ for any $x$, $y$, $f$},
\end{align}
\begin{align}
  D^{[0]}(A, B)
  &\equiv
    \sum_{i a} n_i \bar{n}_a A^{[1]}_{i a} B^{[1]}_{a i}
    + \frac{1}{4} \sum_{i j a b} n_i n_j \bar{n}_a \bar{n}_b A^{[2]}_{i j a b} B^{[2]}_{a b i j},
    \label{eq:flow0} \\
  D^{[1]}_{p, q}(A, B)
  &\equiv
    \sum_{r} A^{[1]}_{p r} B^{[1]}_{r q}
    - \frac{1}{2} \sum_{i j a} n_i n_j \bar{n}_a A^{[2]}_{i j a q} B^{[2]}_{a p i j}
    + \frac{1}{2} \sum_{i a b} n_i \bar{n}_a \bar{n}_b A^{[2]}_{i p a b} B^{[2]}_{a b i q}
    \notag \\
  &\hphantom{=}
    + \sum_{i a} n_i \bar{n}_a \left(
    A^{[1]}_{i a} B^{[2]}_{a p i q}
    + A^{[2]}_{i p a q} B^{[1]}_{a i}
    \right)
    \label{eq:flow1} \\
  D^{[2]}_{p q r s}(A, B)
  &\equiv
    \tilde{\mathcal{A}}_{p, q} \tilde{\mathcal{A}}_{r s}
    \sum_{i a} n_i \bar{n}_a A^{[2]}_{i p a r} B^{[2]}_{a q i s}
    \notag \\
  &\hphantom{=}
    + \frac{1}{2} \sum_{i j} n_i n_j A^{[2]}_{i j r s} B^{[2]}_{p q i j}
    + \frac{1}{2} \sum_{a b} \bar{n}_a \bar{n}_b A^{[2]}_{p q a b} B^{[2]}_{a b r s}
    \notag \\
  &\hphantom{=}
    + \sum_{t} \left(
    \tilde{\mathcal{A}}_{p q} A^{[1]}_{q t} B^{[2]}_{p t r s}
    + \tilde{\mathcal{A}}_{r s} A^{[2]}_{p q r t} B^{[1]}_{t s}.
    \right)
    \label{eq:flow2}
\end{align}

@@

The White generator is explicitly constructed the following way \cite{PhysRevLett.106.222502,White:cond-mat0201346}.  Let
\begin{align*}
\hat{\eta}^{\text{Wh}} &\equiv \hat{\eta}' - \hat{\eta}'{}^\dagger,
\end{align*}
where
\begin{align*}
\eta^{\prime [1]}_{a i} &\equiv \frac{\bar{n}_a n_i H^{[1]}_{a i}}{\tilde{\Delta}^{[1]}_{a i}} &
\eta^{\prime [2]}_{a b i j} &\equiv \frac{\bar{n}_a \bar{n}_b n_i n_j H^{[2]}_{a b i j}}{\tilde{\Delta}^{[2]}_{a b i j}}
\end{align*}
and the Epstein--Nesbet denominators $\tilde{\Delta}$ are defined as {\bf ref to epstein-nesbet deno, insert later}
\begin{align}
\tilde{\Delta}^{[1]}_{a i} &\equiv \Delta_{\{a\}, \{i\}} - H^{[2]}_{a i a i}, \notag \\
\tilde{\Delta}^{[2]}_{a b i j} &\equiv \Delta_{\{a, b\}, \{i, j\}} + w_{a b i j}, \notag \\
\Delta^{[2]}_{P_1 P_2} &\equiv \sum_{p_1 \in P_1} H^{[1]}_{p_1 p_1} - \sum_{p_2 \in P_2} H^{[1]}_{p_2 p_2}, \label{eq:moellerplessetdenominator} \\
w_{a b i j}
  &\equiv H^{[2]}_{a b a b} - H^{[2]}_{a i a i} - H^{[2]}_{b i b i} \notag \\
  &\qquad + H^{[2]}_{i j i j} - H^{[2]}_{a j a j} - H^{[2]}_{b j b j}. \notag
\end{align}
Some authors {\bf ref} use the M\o ller--Plesset denominators $\Delta$ instead, leading to a slightly different variant of the White generator.  {\bf define Moller plesset denos and give reference}.

@@

We can then express the perturbation expansion in terms of summations over matrix elements as we did for the IM-SRG flow equation, again with the aid of Wick's theorem or diagrammatic techniques.  This leads to the following expression for the second-order correction:
\begin{align*}
  \varepsilon_p^{(2)}
  &=
    \sum_{i a b} \frac{n_i \bar{n}_a \bar{n}_b |H^{[2]}_{p i a b}|^2}{2 \Delta_{\{p i\}, \{a b\}}}
    - \sum_{i j a} \frac{n_i n_j \bar{n}_a |H^{[2]}_{i j p a}|^2}{2 \Delta_{\{i j\}, \{p a\}}},
\end{align*}
where the M\o ller--Plesset denominators $\Delta$ are defined in Eq.~\eqref{eq:moellerplessetdenominator}.  The above expression is depicted in Figure \ref{fig:diagrams-sfe}.  Since there are numerous terms in the third-order correction $\varepsilon_p^{(3)}$, they are listed in diagrammatic form in Figure \ref{fig:diagrams-sfe}.  We will refer to QDPT to third order as ``QDPT3''.
